On Math

### 2025-08-29 cards/paper.tex
move statistics to new document, prepare doc for linear algebra also

\section{Statistics}
\subsection{Estimating a mean and variance}
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Reading times (in minutes)} & \textbf{Frequency} \\
    \hline
    $0 \leqslant x < 20$ & $5$ \\
    \hline
    $20 \leqslant x < 40$ & $10$ \\
    \hline
    $40 \leqslant x < 60$ & $8$ \\
    \hline
    $60 \leqslant x < 80$ & $7$ \\
    \hline
  \end{tabular}
  \caption{Estimating a mean and variance}
  \label{tab:estimating_a_mean_and_variance}
\end{table}
To estimate a mean we need a midpoint for $a \leqslant x < b$, so midpoint is $\frac{a+b}{2}$
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{Reading times (in minutes)} & \textbf{Frequency} & \textbf{Midpoint} & \textbf{F*M} \\
    \hline
    $0 \leqslant x < 20$ & $5$ & $10$ & $50$ \\
    \hline
    $20 \leqslant x < 40$ & $10$ & $30$ & $300$ \\
    \hline
    $40 \leqslant x < 60$ & $8$ & $50$ & $400$ \\
    \hline
    $60 \leqslant x < 80$ & $7$ & $70$ & $490$ \\
    \hline
  \end{tabular}
  \caption{Not sure what to type in here}
  \label{tab:estimating_a_mean_and_variance_2}
\end{table}
\begin{equation}
  \overline{\rm x} = \frac{50+300+400+490}{30} = \frac{1240}{30} \approx 41.3
\end{equation}
Mean estimation for continous grouped data:
\begin{equation}
  \overline{\rm x} \approx \frac{1}{n}\Sigma^K_{i=1}(f_i * m_i)
\end{equation}

$f_i$ - frequency of the i-th class

$m_i$ - midpoint of the i-th class

$n$ - total number of the observations

Variance estimation
\begin{equation}
  \sigma^2_n \approx \frac{1}{n}\Sigma^K_{i=1}(f_i * (m_i - \overline{\rm x})^2)
\end{equation}

$f_i$ - frequency of the i-th class

$m_i$ - midpoint of the i-th class

$n$ - total number of the observations

$\overline{\rm x}$ - the mean

\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    \textbf{Midpoint} & \textbf{F*M} & \textbf{$M_i * \overline{\rm x}$} & \textbf{($M_i - \overline{\rm x})^2$} & \textbf{$f_i * (M_i - \overline{\rm x})$} \\
    \hline
    $10$ & $50$ & $-31.333$ & $981.776$ & $4908.879$ \\
    \hline
    $10$ & $30$ & $-11.333$ & $128.444$ & $1284.437$ \\
    \hline
    $50$ & $400$ & $8.667$ & $75.112$ & $600.894$ \\
    \hline
    $70$ & $490$ & $28.667$ & $821.780$ & $5752.458$  \\
    \hline
  \end{tabular}
  \caption{Not sure what to type in here}
  \label{tab:estimating_a_mean_and_variance_3}
\end{table}
\begin{equation}
  \begin{gathered}
    \sigma^2_n \approx \frac{4908... + 1284... + 600... + 5752}{30} \approx 418.2  \\
    \sigma_n \approx \sqrt{418.2} \approx 20.5
  \end{gathered}
\end{equation}
\subsection{Sum of squares}
Given that variance is:
\begin{equation}
  \sigma^2_x = \frac{1}{n} \sum_{i=1}^n(x_i - \overline{\rm x})^2
\end{equation}
The sum of squares is the same but with no fraction:
\begin{equation}
  S_{xx} = \sum_{i=1}^n(x_i - \overline{\rm x})^2
\end{equation}
Compare covariance with a sum of squares:
\begin{equation}
  \begin{gathered}
    Cov(x,y) = \frac{1}{n} \sum_{i=1}^n(x_i - \overline{\rm x})(y_i - \overline{\rm y}) \\
    S_{xy} = \sum_{i=1}^n(x_i - \overline{\rm x})(y_i - \overline{\rm y})
  \end{gathered}
\end{equation}
Variance and covariance in terms of $S_{xx}$, $S_{yy}$ and $S_{xy}$:
\begin{equation}
  \begin{gathered}
    \sigma^2_x = \frac{S_{xx}}{n} \\
    \sigma^2_y = \frac{S_{yy}}{n} \\
    Cov(x,y) = \frac{S_{xy}}{n}
  \end{gathered}
\end{equation}
% TODO: some heading?
Shortcut formulas:
\begin{equation}
  \begin{gathered}
    S_{xx} = \sum_{i=1}^n x_i^2 - \frac{1}{n} (\sum_{i=1}^n x_i)^2 \\
    S_{yy} = \sum_{i=1}^n y_i^2 - \frac{1}{n} (\sum_{i=1}^n y_i)^2 \\
    S_{xy} = \sum_{i=1}^n x_iy_i - \frac{1}{n} (\sum_{i=1}^n x_i)(\sum_{i=1}^n y_i)
  \end{gathered}
\end{equation}
\subsection{Linear regression}
For a set of $n$ paired observations $x_i$ and $y_i$ the trend line is the
straight line that best fits the data.
Finding the equation of the trend line is called \textbf{linear regression}
or best fit.
We can show that the equation of the trend line is given by:
\begin{equation}
  y = mx * b
\end{equation}
1. Slope $m$
\begin{equation}
  m = \frac{S_{xy}}{S_{xx}}
\end{equation}
2. y-intercept - b:
\begin{equation}
  b = \overline{\rm y} - m\overline{\rm x}
\end{equation}
3. $\overline{\rm x}$ and $\overline{\rm y}$ are the means of the data sets:
\begin{equation}
  \overline{\rm x} = \frac{1}{n} \sum_{i=1}^nx_i, \overline{\rm y} = \frac{1}{n} \sum_{i=1}^ny_i,
\end{equation}
4. quantities $S_{xx}$ and $S_{xy}$:
\begin{equation}
  \begin{gathered}
    S_{xx} = \sum_{i=1}^n (x_i - \overline{\rm x})^2 \\
    S_{xy} = \sum_{i=1}^n(x_i - \overline{\rm x})(y_i - \overline{\rm y})
  \end{gathered}
\end{equation}
There is some nice intuition behind the formula for $b$. The trend line
always passes through the point $(\overline{\rm x}, \overline{\rm y})$.
Therefore, this point must satisfy the trend line equation:
\begin{equation}
  y = mx + b
\end{equation}
We can then solve this equation for $b$. Also notice that $S_{xx}$ and $S_{xy}$
are the summation terms that feature in the formulas for the variance and
covariance respectively.
\subsection{Linear coefficient}
Signed with $\rho$ (Greek letter called 'rho').
- $\rho = 1$ means perfect positive correlation
- $\rho = -1$ means perfect negative correlation
- $\rho = 0$ no (linear) correlation
\subsection{Linear correlation coefficient}
$|p| > 0.7$ - correlation is strong
$0.3 \leqslant |p| \leqslant 0.7$ - correlation is weak
$0 \leqslant |p| \leqslant 0.3$ - correlation is negligible
\subsection{Linear correlation coefficient using sigma notation}
\begin{equation}
  p = \frac{S_{xy}}{\sqrt{S_{xx} * S_{yy}}}
\end{equation}
\subsection{Residuals and residual plots}
For dataset of paired numerical observations $(x, y)$ and the corresponding
linear regression model:
\begin{equation}
  \hat y = f(x)
\end{equation}
The \textbf{residual} for a particular observation $(x, y)$ is defined as:
\begin{equation}
  \begin{gathered}
    \text{residual} = \text{actual} - \text{estimated} \\
    y - f(x) = y - \hat y
  \end{gathered}
\end{equation}
We subtract the estimated value predicted by model from the actual observed
value:
\begin{itemize}
\item negative residual - model overestimates the true value
\item positive residual - model underestimates the true value
\end{itemize}

Exercises:
\begin{equation}
  \begin{gathered}
    \hat y = 0.02x + 0.8 \\
    \text{for point} P(400, 8.3) \\
    \hat y = 0.02 * 400 + 0.8 = 8.8 \\
    y - \hat y = 8.3 - 8.8 = -0.5
  \end{gathered}
\end{equation}
\section{Logarithms}
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Property} & \textbf{Formula} \\
    \hline
    Product Rule & $\log_a(xy) = \log_a(x) + \log_a(y)$ \\
    \hline
    Quotient Rule & $\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)$ \\
    \hline
    Power Rule & $\log_a(x^n) = n \cdot \log_a(x)$ \\
    \hline
    Change of Base & $\log_a(x) = \frac{\log_b(x)}{\log_b(a)}$ \\
    \hline
    Equating arguments & $\log_af(x) = log_ag(x)$ is the same as $f(x) = g(x)$ \\
    \hline
  \end{tabular}
  \caption{Fundamental Properties of Logarithms}
  \label{tab:log_properties}
\end{table}

>>>>

\section{Logarithms}
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Property} & \textbf{Formula} \\
    \hline
    Product Rule & $\log_a(xy) = \log_a(x) + \log_a(y)$ \\
    \hline
    Quotient Rule & $\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)$ \\
    \hline
    Power Rule & $\log_a(x^n) = n \cdot \log_a(x)$ \\
    \hline
    Change of Base & $\log_a(x) = \frac{\log_b(x)}{\log_b(a)}$ \\
    \hline
    Equating arguments & $\log_af(x) = log_ag(x)$ is the same as $f(x) = g(x)$ \\
    \hline
  \end{tabular}
  \caption{Fundamental Properties of Logarithms}
  \label{tab:log_properties}
\end{table}
